Integrating and prioritizing ecosystems services at early development stages of industrial systems in a territory: the case of distributed recycling at Nancy, France

### 2022-02-14 docs/PubliER/PubliER.tex
presentation

% Pandoc header
\usepackage{float}
\usepackage{subfig}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage[bitstream-charter]{mathdesign}
\usepackage{pdflscape}
\usepackage{svg}
\usepackage{lineno}
\usepackage{setspace}
\newcommand*{\doverline}[1]{\overline{\overline{#1}}}
\usepackage{tabu}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{hhline}
\usepackage{hyperref}

@@

  \begin{abstract}
  Ecosystem services (ES) is a powerful conceptual framework to put in evidence the benefits that humans receive from nature, most of time for free.
  The problem is that it is complex to link between the local ecosystem services of an urban territory and the industrial systems placed within it, identifying the priority interactions (as synergy and impacts).
  From a decision-maker perspective, there is no a aid decision tool to guide a multicriteria evaluation at early development stages of industrial systems.
  On the ecological side, there have been major efforts in the valuation of ES given by ecosystems (terrestrial, aquatic, atmospheric) to the human well-being and great advances in the developing of a standard ES baselines such as Common International Classification of Ecosystem Services (CICES).
  While on the industrial side, life cycle assessment (LCA) tool is focused on the quantification of the environmental impact of industrial interventions. Despite its ability to cover mutually exclusive and exhaustive impact categories, the LCA approach still has deficiencies: 1) it focuses on quantifying and reducing the net environmental impacts, but not on reducing ecological overshoot and establishing synergies with nature; 2) it considers the interactions between technological processes, but not the interactions between relevant ecosystems.
  Few researches have addressed the alignment the territorial priorities in terms of ES for planning and urban development with the supply/demand of ES by industrial systems.
  There is a need to include the ES in the desicion making process to establish the capacity of the territory in order to evaluate an absolute environmental sustainability of industrial systems.
  The purpose of this article is to propose a methodological approach in order to include ecosystem services in regarding the territorial and industrial endeavors.
  A prioritization is made based on the connection of urban ES services and the techno-ecological synergy.
  The results of are step forwards to create techno-ecological synergies between ecological and industrial systems.
  This methodological steps will applied to the case of distributed recycling via additive manufacturing (DRAM) to highlight the relevant ES from the CICES framework for the territory of Nancy, France.
  The technical advancements of recycling approaches using additive manufacturing are promising technical interventions to foster plastic recycling at a local level.
  \end{abstract}

@@

The current ecological urgency confirms that the understanding and managing the interactions between humans systems and the rest of nature is a major prerequisite for addressing the worsening environmental and social crises of the \(21^{st}\) century (\protect\hyperlink{ref-Lomborg2020}{Lomborg, 2020}).
Even though engineering products and processes might meet the needs of the present society, many of the technological developments are compromising the ability of future generations to meet their own needs (\protect\hyperlink{ref-Bakshi2019a}{Bakshi, 2019}).
For claiming for sustainability in technological development and interventions, the human activities should not exceed critical ecosystem capacity (\protect\hyperlink{ref-Bakshi2018}{Bakshi et al., 2018}).
This meta-principle is a imperative but not sufficient condition for sustainability.
Moreover, no country currently meets minimum thresholds for social development without exceeding planetary boundaries (\protect\hyperlink{ref-ONeill2018}{O'Neill et al., 2018}).
Therefore, it is no possible to rely only on techno-centric interventions without considering the finite planetary ecosystem characterized by profound uncertainty and the shared goals of ecological sustainability and just distribution (\protect\hyperlink{ref-Diwekar2021}{Diwekar et al., 2021}).
We need to integrate ecological carrying capacity since the fuzzy front end phase of an industrial systems.
However, the integration of ecological aspects in the decision-making seems not evident given the complexity to define the boundaries and interactions of industrial and ecological systems.

\protect\hyperlink{ref-Ceschin2016}{Ceschin and Gaziulusoy} (\protect\hyperlink{ref-Ceschin2016}{2016}) putted forward the evolution of \emph{Design for Sustainability (DfS)} framework showing the different approaches that have evolved from a product innovation level to socio-technical systems level.
They pointed out that engineering interventions at only technological unit operation/product level are necessary, but not sufficient condition for sustainability.
\protect\hyperlink{ref-SousaRocha2019}{Rocha et al.} (\protect\hyperlink{ref-SousaRocha2019}{2019}) revealed ten different models at operational, tactical, and strategical levels
One strategical point in sustainability relies on the economic valuation of ecosystem goods and services framework. This approach gives an elegant framework highlighting their importance for society and human welfare.
However, there is a need to explicitly account for their contribution when designing and developing products and services (\protect\hyperlink{ref-Diwekar2021}{Diwekar et al., 2021}).
The engineering discipline developed the implicit assumption that ecological systems have nearly endless capacity to provide resources and adsorb wastes.
This blindness in the engineering vision can be explaining by the fact that at the beginning of the technological industrialization, the human activites' impacts on the earth remained marginal. This scenario is not true today.
The need for ecosystem services research has become evident due to the impacts of population growth, economic activities, and urbanization on natural capital (\protect\hyperlink{ref-Torres2021}{Torres et al., 2021}).
The loss in value associated with biodiversity loss and the related loss of ecosystem services is often invisible and does not influence decision makers (\protect\hyperlink{ref-Bruel2018}{Bruel et al., 2019}).
It is difficult to provide information about pressure from industrial systems in corporate information systems.
Engineering within ecological constraints need to acknowledge the capacity of relevant ecosystems to supply the demanded goods and services while the ecosystems and natural capital must be protected, restored and developed to be capable of continuing to supply those services that industry (and society) relies on (\protect\hyperlink{ref-Bakshi2018}{Bakshi et al., 2018}).

According to the Milleniums 15 out of 24 ecosystems services examined are degraded or being used in an unsustainable manner (\protect\hyperlink{ref-MEA2005}{MEA, 2005}).
Likewise, using the planetary boundaries framework, it is argued that anthropogenic activities already exceed the biophysical limits of the ``safe operating zone'' in terms of carbon and nitrogen cycles, and biodiversity loss (\protect\hyperlink{ref-ONeill2018}{O'Neill et al., 2018}; \protect\hyperlink{ref-Rockstrom2009}{Rockström et al., 2009}).
Among the root causes of ecological degradation is ignorance about the exceed of the ecological carrying capacity in many decisions (\protect\hyperlink{ref-Liu2019g}{Liu and Bakshi, 2019}; \protect\hyperlink{ref-Rodrigues2021}{Rodrigues et al., 2021}).
Another crucial issue is that current design approaches based on life cycle characterization and footprint methods focus on continuous improvement by reducing life cycle impacts per unit of product, encouraging improvements by doing ``less bad,'' which need not translate into keeping human activities within ecological constraints.
Ideally, it is needed to (re)designed industrial activities to reduce the demand for the demand of ecosystems services creating for a local `\emph{island of sustainability}' (\protect\hyperlink{ref-Wallner1996}{Wallner et al., 1996}) which is that the demand should not exceed the supply at the local scale (\protect\hyperlink{ref-Gopalakrishnan2016}{Gopalakrishnan et al., 2016}).
Therefore, it's urgent to expand the boundaries for engineering design from the lowest molecular level to the process level, and from individual process to the higher levels of value chains, ecosystems and the planet (\protect\hyperlink{ref-Martinez-Hernandez2017}{Martinez-Hernandez, 2017}).

The main purpose of this article is to propose a methodology in order to prioritize the ecosystem services at early development stages of industrial systems in a territory.
This methodology is based in the use of the ecosystem services as boundary object (\protect\hyperlink{ref-Abson2014}{Abson et al., 2014}) with the purpose to identify ES according to the perspective of global value chain, industrial chain and territorial perspective.
As a case application, the study of distributed recycling manufacturing will be describe.
Plastic pollution is a global concern that must be addressed collectively with the utmost priority because it endangers the ecosystem and all life forms (\protect\hyperlink{ref-Kumar2021}{Kumar et al., 2021}).
Therefore, new approaches need to be explored in order to reduce or at least recycling this material.
Thus, this study intends to explore the local impact derived from the implementation of a distributed plastic recycling chain in a territory.
The expected results seek to address the following questions:

\begin{itemize}
\tightlist
\item
  What are the appropriate ecosystem service indicators for assessing an prospective filière as distributed plastic recycling? - How does the implementation of a distributed plastic recycling chain impact on ecosystem services?
\item
  What are the barriers and drivers for the development of a distributed plastic recycling chain in a territory?
\end{itemize}

@@

Foundational ideas on ecosystem services (ES) seek for conceptual and methodological tools with the major goal to increase public interest in biodiversity conservation through the recognition, accounting and valuation of the societal dependence on the ecological life support systems for the human well-being (\protect\hyperlink{ref-DeGroot2002}{De Groot et al., 2002}; \protect\hyperlink{ref-Gomez-Baggethun2010}{Gómez-Baggethun et al., 2010}).
ES describes the ecological characteristics, function or processes that contribute (actively or passively) to the human well-being (\protect\hyperlink{ref-Costanza2017}{Costanza et al., 2017}, \protect\hyperlink{ref-Costanza1997}{1997}).
Ecosystem goods (e.g; Food) and services (e.g.~waste assimilation) illustrate the benefits that human derive from the ecosystem functions (\protect\hyperlink{ref-Costanza1997}{Costanza et al., 1997}).
It is needed to distinguish between the ecosystem's functions and processes from the ecosystem services concept itself.
The former describes biophysical relationships that are carried out by nature regardless of whether or not human benefits.
By contrast, the latter are those processes and functions where people can (or could have the potential) obtain benefits.
The ecosystem services do not flow to human well-being without crucial interactions with the different forms of capital (Natural, Social, Human, Built), which entails the need of understanding, modelling, measuring, and managing ES in a trans-disciplinary approach.
Likewise, the concept of ecosystem dis-service denotes the processes and functions that affect humans in `negative' way, making damage and costs (\protect\hyperlink{ref-TEEB2010}{TEEB, 2010}).
One major point that ES make clear is to raise awareness on the recognition of humanity's primary dependencies on the `functions of' natural capital which reflects the fact that, however they may perceive themselves, humans are part of, and not apart from, nature (\protect\hyperlink{ref-Ekins2003}{Ekins et al., 2003}).
This entails the necessity to create knowledge for trans-disciplinary approaches using ES as boundary object for sustainability for diverse stakeholders (\protect\hyperlink{ref-Honeck2021}{Honeck et al., 2021}).

@@

Efforts on biodiversity conservation relies on the highlight of the economic aspects of biodiversity and the natural capital (\protect\hyperlink{ref-Costanza1997}{Costanza et al., 1997}) and the environmental inaction related to the cost of policy damage occurring in the absence of an effective regulatory framework (\protect\hyperlink{ref-Bruel2016}{Bruel et al., 2016}).
From a strong sustainability perspective, a declining capital stock is an unambiguous indicator of unsustainability in the flow of goods and services that derive from it (\protect\hyperlink{ref-Ekins2003}{Ekins et al., 2003}).
More important, the recognition of the non-substitutability of natural capital with regard to the other forms of capital; acknowledging the characteristics of irreversibility (such as species extinction or climate change), uncertainty and the existence of \emph{critical} components that make a major contribution to welfare.
The main core of the environmental problem relies on the use of use ecosystem's functions, mainly those that generate economic welfare, that are making a negative impact and influence on the natural capital stock, and even worse, on those functions that are responsible for ecosystem stability and resilience (\protect\hyperlink{ref-Ekins2003}{Ekins et al., 2003}).

@@

Different initiatives have been reported to classify the ES, including the Millennium Ecosystem Assessment (\protect\hyperlink{ref-MEA2005}{MEA, 2005}),
The Economics of Ecosystems and Biodiversity TEEB (\protect\hyperlink{ref-TEEB2010}{TEEB, 2010}),
The Intergovernmetal Plastform of Biodiversity and Ecosystem Services (IPBES) (\protect\hyperlink{ref-ref}{\textbf{ref?}}) and the Common International Classification of Ecosystem Services (CICES).
In the heart of the four main, they share four main categories of ES:
\textbf{Provisioning} (e.g.~food and medicines);
\textbf{Regulating} (e.g.~pollination and climate regulation),
\textbf{Supporting} (e.g soil formation and fixation of solar energy) and
\textbf{Cultural / Information} services (e.g.~artistic inspiration and recreation) services are four broad categories types of ES constitutes the core of most recent classifications and that are shared by the most frameworks (\protect\hyperlink{ref-PedersenZari2019}{Pedersen Zari, 2019}).

@@

{\centering \includegraphics[width=1\linewidth]{Figures/Comparison}

@@

The Common International Classification of Ecosystem Services (CICES) was developed to provide hiecharchally consistent and science-based classification to be used for natural capital accounting purposes (\protect\hyperlink{ref-ref}{\textbf{ref?}}).
In CICES framework, \protect\hyperlink{ref-Potschin-Young2018}{Potschin-Young et al.} (\protect\hyperlink{ref-Potschin-Young2018}{2018}) argued the conceptual framework of cascading aspect from ecosystesm service are commonly divided Groups, division.

@@

{\centering \includegraphics[width=1\linewidth]{Figures/cices}

@@

\caption{ES conceptual framework}\label{fig:Fig-ES-cices}
\end{figure}

The CICES scheme, on the other hand, refers to ES specifically as the products from ecosystem, which can be directly used by human-beings {[}70{]}.
With this definition, supporting services defined in MEA scheme are excluded from CICES scheme, because supporting services cannot be directly consumed by human beings.
The product-based CICES framework can enable clearer mapping between ecological and economic products.
The hierarchical structure of CICES scheme has provisioning, regulating and cultural services at the highest (section) level.

The following levels organize ES according to their divisions, groups and classes.
At lower levels, more specific services can be defined and characterized.

For example, at section level, we may have only the provisioning service but at class level, we may define the service to be forage grass cultivated as animal feed.

Therefore, the hierarchical structure can supply different levels of detail needed by decision makers.

\hypertarget{ecosystems-services-and-industrial-systems-towards-a-synergy-between}{%
\subsection{Ecosystems services and industrial systems: Towards a synergy between}\label{ecosystems-services-and-industrial-systems-towards-a-synergy-between}}

One major point A holistic evaluation of the externalities that industrial system
Ecosystems services and industrial systems need to be confronten to holistic evaluation
The key point for bussiness and government sectors in terms realize an integrated valuation and comprehensive account of both the negative impacts on ES from business activites and the positive contributions of ES to business and households (\protect\hyperlink{ref-Costanza2017}{Costanza et al., 2017}).

\protect\hyperlink{ref-Bakshi2015}{Bakshi et al.} (\protect\hyperlink{ref-Bakshi2015}{2015}) reported a framework of Techno-Ecological Synergy (TES) in order to expand the scope of the usual techno-centric perspectives.
The main point argued is that TES develops ways of enhancing synergies between a local scale manufacturing process and the land around it.
The final aims is to encourage a more robust analysis of technological and ecological systems at multiple spatial scales ranging from local (e.g.~for small systems such as a house and its yard, a manufacturing process and its site) to a larger scale systems that extend to consider the entire life cycle.

@@

More than two thirds of the world's population are expected to live in cities by 2050 (\protect\hyperlink{ref-ref}{\textbf{ref?}}).
Thus, the integration of ecosystem services in the urban and peri-urban planning is a key issue for on quality of life in cities (\protect\hyperlink{ref-Gomez-Baggethun2013}{Gómez-Baggethun and Barton, 2013}).
Urban ecosystems are especially important in providing services with direct impact on health and security such as air purification, noise reduction, urban
cooling, and runoff mitigation (\protect\hyperlink{ref-Bolund1999}{Bolund and Hunhammar, 1999}).
\protect\hyperlink{ref-Gomez-Baggethun2013}{Gómez-Baggethun and Barton} (\protect\hyperlink{ref-Gomez-Baggethun2013}{2013}) draws on the developments of the ecosystem services for urban planning.
They formulated a classification of ecosystem services and disservices in urban areas, including valuation methods and tools.

@@

The purpose of this article is to identify main relevant ES services from a industrial and territorial perspective at the early design of industrial systems.
This identification is a first step with the purpose to evaluate the supply and demand of ecosystem services for these industrial systems.
A prospective distributed recycling chain for additive manufacturing.
A conceptual framework to evaluate the synergy of prospective industrial filières considering the technological and ecological spheres.

@@

{\centering \includegraphics[width=1\linewidth]{Figures/Methodology}

@@

@@

\renewcommand{\arraystretch}{1.5}
\begin{table}[!h]

>>>>

% Pandoc header
\usepackage{float}
\usepackage{subfig}
\usepackage[utf8]{inputenc}
\def\tightlist{}
\usepackage[bitstream-charter]{mathdesign}
\usepackage{pdflscape}
\usepackage{svg}
\usepackage{lineno}
\usepackage{setspace}
\newcommand*{\doverline}[1]{\overline{\overline{#1}}}
\usepackage{tabu}
\makeatletter
\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
         {12\p@ \@plus 6\p@ \@minus 3\p@}%
         {3\p@ \@plus 6\p@ \@minus 3\p@}%
         {\normalfont\normalsize\itshape\bfseries}}
 \makeatother        
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

  \begin{abstract}
  Ecosystem services (ES) is a powerful conceptual framework to put in evidence the benefits that humans receive from nature, most of time for free.
  From a decision-maker perspective, there is no a aid decision tool to guide a multicriteria evaluation at early development stages of industrial systems.
  On the ecological side, there have been major efforts in the valuation of ES given by ecosystems (terrestrial, aquatic, atmospheric) to the human well-being and great advances in the developing of a standard ES baselines such as Common International Classification of Ecosystem Services (CICES).
  While on the industrial side, life cycle assessment (LCA) tool is focused on the quantification of the environmental impact of industrial interventions. Despite its ability to cover mutually exclusive and exhaustive impact categories, the LCA approach still has deficiencies: 1) it focuses on quantifying and reducing the net environmental impacts, but not on reducing ecological overshoot and establishing synergies with nature; 2) it considers the interactions between technological processes, but not the interactions between relevant ecosystems.
  Few researches have addressed the alignment the territorial priorities in terms of ES for planning and urban development with the supply/demand of ES by industrial systems.
  There is a need to include the ES in the desicion making process to establish the capacity of the territory in order to evaluate an absolute environmental sustainability of industrial systems.
  The purpose of this article is to propose a methodological approach in order to include ecosystem services in regarding the territorial and industrial endeavors.
  A prioritization is made based on the connection of urban ES services and the techno-ecological synergy.
  The results of are step forwards to create techno-ecological synergies between ecological and industrial systems.
  This methodological steps will applied to the case of distributed recycling via additive manufacturing (DRAM) to highlight the relevant ES from the CICES framework for the territory of Nancy, France.
  The technical advancements of recycling approaches using additive manufacturing are promising technical interventions to foster plastic recycling at a local level.
  \end{abstract}

@@

The current ecological urgency confirms that the understanding and managing the interactions between humans systems and the rest of nature is a major prerequisite for addressing the worsening environmental and social crises of the \(21^{st}\) century (\protect\hyperlink{ref-Lomborg2020}{Lomborg, 2020}).
Even though engineering products and processes might meet the needs of the present society, many of the technological developments are compromising the ability of future generations to meet their own needs (\protect\hyperlink{ref-Bakshi2019a}{Bakshi, 2019}).
Claiming for sustainability in technological development and interventions, implies that the human activities should not exceed critical ecosystem capacity (\protect\hyperlink{ref-Bakshi2018}{Bakshi et al., 2018}).
This meta-principle is an imperative but not sufficient condition for sustainability.
Moreover, no country currently meets minimum thresholds for social development without exceeding planetary boundaries (\protect\hyperlink{ref-ONeill2018}{O'Neill et al., 2018}).
According to the Milleniums 15 out of 24 ecosystems services examined are degraded or being used in an unsustainable manner (\protect\hyperlink{ref-MEA2005}{MEA, 2005}).
Likewise, using the planetary boundaries framework, it is argued that anthropogenic activities already exceed the biophysical limits of the ``safe operating zone'' in terms of carbon and nitrogen cycles, and biodiversity loss (\protect\hyperlink{ref-ONeill2018}{O'Neill et al., 2018}; \protect\hyperlink{ref-Rockstrom2009}{Rockström et al., 2009}).
Among the root causes of ecological degradation is ignorance about the exceed of the ecological carrying capacity in many decisions (\protect\hyperlink{ref-Liu2019g}{Liu and Bakshi, 2019}; \protect\hyperlink{ref-Rodrigues2021}{Rodrigues et al., 2021}).
Therefore, it is no possible to rely only on techno-centric interventions without considering the finite planetary ecosystem characterized by profound uncertainty and the shared goals of ecological sustainability and just distribution (\protect\hyperlink{ref-Diwekar2021}{Diwekar et al., 2021}).
It is urgent to expand the boundaries for engineering design from the lowest molecular level to the process level, and from individual process to the higher levels of value chains, ecosystems and the planet (\protect\hyperlink{ref-Martinez-Hernandez2017}{Martinez-Hernandez, 2017}).
We need to integrate ecological carrying capacity since the fuzzy front end phase of an industrial systems.
However, the integration of ecological aspects in the decision-making seems not evident given the complexity to define the boundaries and interactions of industrial and ecological systems.

\protect\hyperlink{ref-Ceschin2016}{Ceschin and Gaziulusoy} (\protect\hyperlink{ref-Ceschin2016}{2016}) putted forward the evolution of \emph{Design for Sustainability (DfS)} framework showing the different approaches that have evolved from a product innovation level to socio-technical systems level.
They pointed out that engineering interventions at only technological unit operation/product level are necessary, but not sufficient condition for sustainability.
\protect\hyperlink{ref-SousaRocha2019}{Rocha et al.} (\protect\hyperlink{ref-SousaRocha2019}{2019}) revealed ten different models at operational, tactical, and strategical levels.
One strategical point in sustainability relies on the economic valuation of ecosystem goods and services framework.
This approach gives an elegant framework highlighting their importance for society and human welfare.
However, there is a need to explicitly account for their contribution when designing and developing products and services (\protect\hyperlink{ref-Diwekar2021}{Diwekar et al., 2021}).
The engineering discipline developed the implicit assumption that ecological systems have nearly endless capacity to provide resources and adsorb wastes.
This blindness in the engineering vision can be explaining by the fact that at the beginning of the technological industrialization, the human activites' impacts on the earth remained marginal. This scenario is not true today.
The need for ecosystem services research has become evident due to the impacts of population growth, economic activities, and urbanization on natural capital (\protect\hyperlink{ref-Torres2021}{Torres et al., 2021}).
The loss in value associated with biodiversity loss and the related loss of ecosystem services is often invisible and does not influence decision makers (\protect\hyperlink{ref-Bruel2018}{Bruel et al., 2019}).
It is difficult to provide information about pressure from industrial systems in corporate information systems.
Engineering within ecological constraints need to acknowledge the capacity of relevant ecosystems to supply the demanded goods and services while the ecosystems and natural capital must be protected, restored and developed to be capable of continuing to supply those services that industry (and society) relies on (\protect\hyperlink{ref-Bakshi2018}{Bakshi et al., 2018}).

The main purpose of this article is to propose a methodology in order to prioritize the ecosystem services at early development stages of industrial systems in a territory.
This methodology is based in the use of the ecosystem services as boundary object (\protect\hyperlink{ref-Abson2014}{Abson et al., 2014}) with the purpose to identify ES according to the perspective of global value chain, industrial chain and territorial perspective.
As a case application, the study of distributed recycling manufacturing will be described.
Plastic pollution is a global concern that must be addressed collectively with the utmost priority because it endangers the ecosystem and all life forms (\protect\hyperlink{ref-Kumar2021}{Kumar et al., 2021}).
Therefore, new approaches need to be explored in order to reduce or at least recycling this material.
Thus, this study intends to explore the local impact derived from the implementation of a distributed plastic recycling chain in a territory.
The expected results seek to address the following questions:

\begin{itemize}
\tightlist
\item
  What are the appropriate ecosystem service indicators for assessing an prospective filière as distributed plastic recycling?
\item
  How does the implementation of a distributed plastic recycling chain impact on ecosystem services?
\item
  What are the barriers and drivers for the development of a distributed plastic recycling chain in a territory?
\end{itemize}

@@

Foundational ideas on ecosystem services (ES) seek for conceptual and methodological tools with the major goal to increase public interest in biodiversity conservation through the recognition, accounting and valuation of the societal dependence on the ecological life support systems for the human well-being (\protect\hyperlink{ref-DeGroot2002}{De Groot et al., 2002}; \protect\hyperlink{ref-Gomez-Baggethun2010}{Gómez-Baggethun et al., 2010}).
ES describes the ecological characteristics, function or processes that contribute (actively or passively) to the human well-being (\protect\hyperlink{ref-Costanza2017}{Costanza et al., 2017}, \protect\hyperlink{ref-Costanza1997}{1997}).
Ecosystem goods (e.g; Food) and services (e.g.~waste assimilation) illustrate the benefits that human derive from the ecosystem functions (\protect\hyperlink{ref-Costanza1997}{Costanza et al., 1997}).
It is needed to distinguish between the ecosystem's functions and processes from the ecosystem services concept itself.
The former describes biophysical relationships that are carried out by nature regardless of whether or not human benefits.
By contrast, the latter are those processes and functions where people can (or could have the potential) obtain benefits.
The ecosystem services do not flow to human well-being without crucial interactions with the different forms of capital (Natural, Social, Human, Built), which entails the need of understanding, modelling, measuring, and managing ES in a trans-disciplinary approach.
Likewise, the concept of ecosystem dis-service denotes the processes and functions that affect humans in `negative' way, making damage and costs (\protect\hyperlink{ref-TEEB2010}{TEEB, 2010}).
One major point that ES make clear is to raise awareness on the recognition of humanity's primary dependencies on the `functions of' natural capital. Humans are part of, and not apart from, nature (\protect\hyperlink{ref-Ekins2003}{Ekins et al., 2003}).
This entails the necessity to create knowledge for trans-disciplinary approaches using ES as boundary object for sustainability for diverse stakeholders (\protect\hyperlink{ref-Honeck2021}{Honeck et al., 2021}).

@@

Therefore, it is needed the correlation of ES with the global value chain and the natural capital.
Efforts on biodiversity conservation relies on the highlight of the economic aspects of biodiversity and the natural capital (\protect\hyperlink{ref-Costanza1997}{Costanza et al., 1997}) and the environmental inaction related to the cost of policy damage occurring in the absence of an effective regulatory framework (\protect\hyperlink{ref-Bruel2016}{Bruel et al., 2016}).
From a strong sustainability perspective, a declining capital stock is an unambiguous indicator of unsustainability in the flow of goods and services that derive from it (\protect\hyperlink{ref-Ekins2003}{Ekins et al., 2003}).
More important, the recognition of the non-substitutability of natural capital with regard to the other forms of capital; acknowledging the characteristics of irreversibility (such as species extinction or climate change), uncertainty and the existence of \emph{critical} components that make a major contribution to welfare.
The main core of the environmental problem relies on the use of use ecosystem's functions, mainly those that generate economic welfare, that are making a negative impact and influence on the natural capital stock, and even worse, on those functions that are responsible for ecosystem stability and resilience (\protect\hyperlink{ref-Ekins2003}{Ekins et al., 2003}).

@@

Different initiatives have been reported to classify the ES, including the Millennium Ecosystem Assessment (\protect\hyperlink{ref-MEA2005}{MEA, 2005}),
The Economics of Ecosystems and Biodiversity TEEB (\protect\hyperlink{ref-TEEB2010}{TEEB, 2010}),
The Intergovernmetal Plateform of Biodiversity and Ecosystem Services (IPBES) (\protect\hyperlink{ref-IPBS2019}{IPBS, 2019}) and the Common International Classification of Ecosystem Services (CICES) (\protect\hyperlink{ref-RoyHaines-Young2018}{Roy Haines-Young and Potschin, 2018}).
Essentially, there are four main categories:
\textbf{Provisioning} (e.g.~food and medicines);
\textbf{Regulating} (e.g.~pollination and climate regulation),
\textbf{Supporting} (e.g soil formation and fixation of solar energy) and
\textbf{Cultural / Information} services (e.g.~artistic inspiration and recreation) services.
These four broad categories constitutes the core of most recent classifications and that are shared by the most frameworks (\protect\hyperlink{ref-Costanza2017}{Costanza et al., 2017}).

@@

{\centering \includegraphics[width=1\linewidth]{/Users/fabio/Documents/4-Projects/everest-bio/PubliER/Figures/Potschin-Burkhard2017}

@@

In the context of this article, the CICES approach is chosen by the fact that it establish hierarchically consistent and science-based classification to be used for natural capital accounting purposes.
\protect\hyperlink{ref-Potschin-Young2018}{Potschin-Young et al.} (\protect\hyperlink{ref-Potschin-Young2018}{2018}) argued the conceptual framework of `ecosystem service cascade' feature make it possible to re-frame the analysis of ES helping in the clarification of the complex relationships that can be placed.
CICES scheme refers to ES specifically as the products from ecosystem, which can be directly used by human-beings.
This implies that \emph{supporting services}, defined in MEA framework, are excluded from CICES scheme, by the fact that human beings do not `receive'consumed' directly the benefits of these services.
Additionally, it tries to avoid the double-counting problem.
In that sense, CICES has \emph{regulation and maintenance services} as one section only.
The other two are \emph{provisioning services} and \emph{cultural services}.
From there, there are \emph{Divisions} describing the main types of output or process, and these are further split into \emph{groups} according to biological, physical or cultural type or process.
The lowest level is the \emph{class}, which specify the biological / material outputs and bio-physical and cultural processes.
Finally, the \emph{class type} measures the \emph{class level} which are the individual entities of the ESs.
Nevertheless, this cascade aspect is not far from critics.
(\protect\hyperlink{ref-Constanza2017}{\textbf{Constanza2017?}}) raised the point that the complex connections among the different ES and the benefits might be poorly represented by a linear `cascade' model, which assumes simple linkages and effects.
But on the other hand, the hierarchical structure can add different levels of detail needed by decision makers.

Figure \ref{fig:ES-CICES} presents the whole structure of ES from the CICES V5.1 {[}@{]}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/fabio/Documents/4-Projects/everest-bio/PubliER/Figures/CICES-diagram}

@@

\caption{ES conceptual framework}\label{fig:ES-CICES}
\end{figure}

\hypertarget{ecosystems-services-and-industrial-systems-towards-a-synergy-between-both}{%
\subsection{Ecosystems services and industrial systems: Towards a synergy between both}\label{ecosystems-services-and-industrial-systems-towards-a-synergy-between-both}}

In environmental management, major efforts have been dedicated to the evaluation of the environmental impact of industrial systems using Life cycle assessment (LCA).
LCA aims to assess impacts on ecosystems, natural resources, and human health, as three primary areas of protection.
LCA proffers an internationally standardized approach to modeling, assessing, and valuing the impacts of a product or process throughout its life cycle.
However, a design approaches based on life cycle characterization and footprint methods focus on continuous improvement by reducing life cycle impacts per unit of product, encouraging improvements by doing ``less bad,'' which need not necessary translate into keeping human activities within ecological constraints.
Moreover, LCA does not fully acknowledge impacts on the supply of ES.
Assessing the impacts of production systems on ES supply is a new -- and contested -- paradigm of the latest years (\protect\hyperlink{ref-Vanderwilde2021}{Vanderwilde and Newell, 2021})
Ideally, it is needed to (re)designed industrial activities to reduce the demand of ecosystems services creating for a local `\emph{island of sustainability}' (\protect\hyperlink{ref-Wallner1996}{Wallner et al., 1996}) which is that the demand should not exceed the supply at the local scale (\protect\hyperlink{ref-Gopalakrishnan2016}{Gopalakrishnan et al., 2016}).

Ecosystems services and industrial systems need to be confronted to holistic evaluation to determine the externalities positive and negative.
In particular, full cost accounting in the business and governmental sectors, including new comprehensive accounts that include both the negative impacts on ES from business activities and the positive contributions of ES to businesses and households are key to develop (\protect\hyperlink{ref-Costanza2017}{Costanza et al., 2017})
Decision makers at the early stages should take into account the potential synergies and threats that the development of industrial systems could have to the local ecosystems.
For instance, \protect\hyperlink{ref-Bakshi2015}{Bakshi et al.} (\protect\hyperlink{ref-Bakshi2015}{2015}) reported a framework of Techno-Ecological Synergy (TES) in order to expand the scope of the usual techno-centric perspectives.
The main point argued is that TES develops ways of enhancing synergies between a local scale manufacturing process and the land around it.
The final aims is to encourage a more robust analysis of technological and ecological systems at multiple spatial scales ranging from local (e.g.~for small systems such as a house and its yard, a manufacturing process and its site) to a larger scale systems that extend to consider the entire life cycle.

In that sense, \protect\hyperlink{ref-Vanderwilde2021}{Vanderwilde and Newell} (\protect\hyperlink{ref-Vanderwilde2021}{2021}) reported a systematic literature review revealing that a persistent division between the LCA and ES research communities.
They putted forward that there are LCA research has focus on one or a few ES, especially carbon balance, sequestration, and emissions.
Also, the work has prioritized regulation and maintenance and provisioning services, with work especially scarce on cultural services.
Likewise, \protect\hyperlink{ref-Rugani2019}{Rugani et al.} (\protect\hyperlink{ref-Rugani2019}{2019}) proposed an approach towards consensus building in the assessment of human activities' impacts on the provision of ES and their beneficial uses for human well-being. More specifically, to harmonize the cascade concept (\protect\hyperlink{ref-RoyHaines-Young2018}{Roy Haines-Young and Potschin, 2018}) from the ES literature and the cause- effect chain modelling used in LCIA.

@@

More than two thirds of the world's population are expected to live in cities by 2050 (\protect\hyperlink{ref-ref}{\textbf{ref?}}).
Thus, the integration of ecosystem services in the urban and peri-urban planning is a key issue for prioritizing the quality of life in cities (\protect\hyperlink{ref-Gomez-Baggethun2013}{Gómez-Baggethun and Barton, 2013}), helping to urban planners to guarantee that future city districts the local people have access to urban greenery and their services

Urban ecosystems are especially important in providing services with direct impact on health and security such as air purification, noise reduction, urban cooling, and runoff mitigation (\protect\hyperlink{ref-Bolund1999}{Bolund and Hunhammar, 1999}).
\protect\hyperlink{ref-Croci2021}{Croci et al.} (\protect\hyperlink{ref-Croci2021}{2021}) presented a critical review
\protect\hyperlink{ref-Gomez-Baggethun2013}{Gómez-Baggethun and Barton} (\protect\hyperlink{ref-Gomez-Baggethun2013}{2013}) draws on the developments of the ecosystem services for urban planning.
They formulated a classification of ecosystem services and disservices in urban areas, including valuation methods and tools.

Using this background on ES from the different perspectives, in the following section a methodology is proposed converge main ES that are relevant for the industrial, territorial and global value chain.

@@

The purpose of this article is to identify main relevant ES services from a industrial and territorial perspective at the early design of industrial systems.
This identification is a first step with the purpose to evaluate the supply and demand of ecosystem services for these industrial systems.

@@

{\centering \includegraphics[width=1\linewidth]{/Users/fabio/Documents/4-Projects/everest-bio/PubliER/Figures/Methodology}

@@

A prospective distributed recycling chain for additive manufacturing .

@@

\renewcommand{\arraystretch}{1.5}
